---
title: "TidyTuesday 2020/42 - The Datasaurus Dozen by Alberto Cairo"
author: "Cedric Scherer"
date: "13th of October 2020"
output:
  html_document:
  theme: paper
highlight: kate
editor_options:
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.showtext = T, fig.retina = 1)
```

```{r prep, message=FALSE, warning=FALSE}
## packages
library(tidyverse)
library(gganimate)
library(ragg)
library(ggtext)
library(colorspace)
library(glue)

theme_set(theme_classic(base_family = "Zilla Slab", base_size = 25))

theme_update(
  axis.text = element_text(family = "Overpass Mono",
                           color = "grey60", margin = margin(t = 2)),
  axis.ticks = element_line(color = "grey60"),
  axis.line.x = element_line(color = "grey60"),
  axis.line.y = element_line(color = "grey60"),
  plot.title = element_text(family = "Zilla Slab Highlight", size = 84, 
                            color = "grey10", hjust = .5,
                            margin = margin(15, 0, 30, 0)),
  plot.subtitle = element_markdown(size = 20, color = "grey30", lineheight = 1.6,
                                   hjust = .5, margin = margin(20, 0, 100, 0)),
  plot.caption = element_text(color = "grey30", size = 18, face = "italic", 
                              lineheight = 1.3,
                              hjust = .5, margin = margin(60, 0, 0, 0)),
  plot.title.position = "plot",
  plot.caption.position = "plot",
  plot.margin = margin(90, 70, 90, 70),
  plot.background = element_rect(color = "white", fill = "white")
)
```

```{r data}
df_saurus <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-13/datasaurus.csv')
```

## Idea and sources

* DataSaurus - Blog Post by Alberto Cairo: thefunctionalart.com/2016/08/download-datasaurus-never-trust-summary.html
* Datasaurus Dozen - Paper by Matejka & Fitzmaurice: autodeskresearch.com/publications/samestats
* Data for R - package by Steph Locke: itsalocke.com/datasaurus/ and github.com/lockedata/datasauRus
* gganimate code by Tom Westlake: r-mageddon.netlify.com/post/reanimating-the-datasaurus
* Based on "Anscombe's Quartet": en.wikipedia.org/wiki/Anscombe%27s_quartet

```{r data-prep}
cols <- c(rcartocolor::carto_pal(12, name = "Prism"), "black")
#cols <- rainbow(13)

df_saurus_lab <- 
  df_saurus %>% 
  group_by(dataset) %>% 
  ## old code that I would now write with `summarize(across())`
  mutate(
    id = cur_group_id(),
    mean_x = mean(x),
    mean_y = mean(y),
    sd_x = sd(x),
    sd_y = sd(y),
    cor = cor(x, y, method = "pearson"),
    label = glue::glue('
      mean(x)  = {formatC(mean_x, digits = 5, format = "f")}
      sd(x)    = {formatC(sd_x, digits = 5, format = "f")}
      
      mean(y)  = {formatC(mean_y, digits = 5, format = "f")}
      sd(y)    = {formatC(sd_y, digits = 5, format = "f")}
      
      cor(x,y) = {formatC(cor, digits = 5, format = "f")}
    '),
    label_short = glue::glue('
      mean(x)  = {floor(mean_x*100)/100}
      sd(x)    = {floor((sd_x + 0.00001)*100)/100}
      
      mean(y)  = {floor(mean_y*100)/100}
      sd(y)    = {floor((sd_y + 0.00001)*100)/100}
      
      cor(x,y) = {ceiling(cor*100)/100}
    ')
  ) %>%
  nest() %>% 
  mutate(modelout = map(data, ~ lm(y ~ x, data = .x) %>%
    broom::tidy() %>% 
    mutate(term = if_else(term == "(Intercept)", "intercept", "slope")) %>% 
    dplyr::select(term, estimate) %>% 
    pivot_wider(
      names_from = term,
      values_from = estimate
    ))) %>%
  unnest() %>% 
  ungroup()
```


```{r plot}
sauR <- 
  ggplot(data = df_saurus_lab,
         aes(x = x, y = y,
             color = dataset, 
             group = 1)) +
    geom_abline(data = df_saurus_lab %>% group_by(dataset) %>% slice(1),
                aes(slope = slope, intercept = intercept, 
                    group = 1, color = dataset, 
                    color = after_scale(lighten(color, .5))), 
                inherit.aes = FALSE,
                size = 1.5) +  
    geom_point(size = 5, alpha = 0.8) +
    geom_text(data = df_saurus_lab %>% group_by(dataset) %>% slice(1),
              aes(x = 92, y = 100, label = label),
              family = "Overpass Mono", color = "grey78", 
              size = 9, hjust = 0, vjust = 1) +
    geom_text(data = df_saurus_lab %>% group_by(dataset) %>% slice(1),
              aes(x = 92, y = 100, label = label_short),
              family = "Overpass Mono", color = "grey60", 
              size = 9, hjust = 0, vjust = 1) +
    coord_equal(clip = "off") +
    scale_x_continuous(limits = c(0, 125), breaks = seq(0, 100, by = 20), expand = c(.01, .01)) +
    scale_y_continuous(limits = c(0, 101), breaks = seq(0, 100, by = 20), expand = c(.01, .01)) +
    scale_color_manual(values = cols, guide = "none") +
    labs(x = NULL, y = NULL, 
         title = "Same Stats. Different Graphs.",
         subtitle = "<b style='font-size:44pt;'>The Famous Datasaurus (*Anscombosaurus spec.*) Dozen</b><br><br>is a set of 13 different datasets with nigh-identical summary statistics, which could lead one to believe the<br>datasets are quite similar. After visualizing the data, it becomes clear that the datasets are markedly different.<br>It is based on **Anscombe's Quartet** which was developed by F.J. Anscombe in 1973 to demonstrate the importance<br>of data visualization. In 2016, Alberto Cairo created the Datasaurus dataset which urges people to *never trust*<br>*summary statistics alone; always visualize your data*. This dataset was published together with a dozen<br> other datasets with almost the same summary statistics by Justin Matejka & George Fitzmaurice in 2017.",
         caption = "Idea by Francis Anscombe, Alberto Cairo, Justin Matejka & George Fitzmaurice\nVisualization by Cédric Scherer based on code by Tom Westlake") + 
    transition_states(dataset, 10, 3) + 
    ease_aes('cubic-in-out')
```

```{r animation}
## png uoutput has weird spacing (problems ggtext + custom fonts), no problem with jpeg
## gifski cannot handle jpeg, magick never finished (10min+)
## solution for now: mpeg video / raw plots and convert manually to gif
anim <- 
  animate(sauR, nframes = 100, fps = 5, detail = 5, 
          width = 1650, height = 2000, device = "jpeg", type = "cairo",
          renderer = ffmpeg_renderer())
          #renderer = magick_renderer())
          #renderer = file_renderer(dir = here::here("plots", "2020_42", "jpeg"), 
          #                         prefix = "2020_42_Datasaurus"))

anim_save(filename = "2020_42_Datasaurus.mpeg", 
          path = here::here("plots", "2020_42"))
```

***

```{r session}
Sys.time()
git2r::repository()
sessionInfo()
```
